### 2014-04-01 paper/designspace.tex
shortening

\section{The Design Space of Pipes}
\bjoern{Budget at most 1.5 columns for this section

 There still some redundancy with the related work section, too}
By using different openings, designs, and topologies of pipes, and inserting different media into the completed pipes, inputs and outputs can be located at arbitrary points on the surface of a 3D print.  We describe the space below.

@@

\tovi{this seems like a forced and
unnecessary definition. Do we really
need to distibguish the two ends?}
Pipes can have four opening styles: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user; and a ``system side"---where other hardware such as electronics may be connected.  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' or ``open on the system side'' to have its contents \emph{physically accessible} to an interacting user or the system \emph{at interaction time}.  Similarly, pipes ``closed on the user side'' or ``closed on the system side'' are physically cut off from the user or machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents.

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/types.png}
\caption{Types of pipes.  A shows a pair of open tubes filled with copper paint to power an LED.  B shows a return pipe with EL wire threaded through it.  C shows two semi-closed pipes with rubber membranes at their ends.  In D, a rabbit has a fully enclosed chamber to modify its weight. \bjoern{Great, but make figure  full column wide}}
\label{fig:pipespace}
\end{figure}

Open pipes originate from the system side and connect the user side, with both ends of the pipe open. This type of pipe may be used to create electronic circuitry.  An open pipe filled with conductive paint can power an LED (see Figure \ref{fig:pipespace}A).

A return pipe originates at the system and returns back to it, without offering physical access to the user.  By threading an electroluminescent (EL) wire through a clear return pipe, a maker can create a custom piece of neon art (see Figure \ref{fig:pipespace}B).

Semi-closed pipes are open at the system end and closed at the user end.  The closed interface may be fabricated from a malleable material: for example, a series of pipes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback (see Figure \ref{fig:pipespace}C).

A fully enclosed pipe is closed to both the system and the user.  Fully enclosed pipes may be used as weight modifiers (e.g., for object identification), as in Figure \ref{fig:pipespace}D).  They can also be used as containers for water or other particles that would otherwise fall out.

\subsection{Design of Pipes}
Pipe design may emphasize either exterior connection points or interior paths.  These two foci lead to different kinds of interfaces.  An example focused on the exterior features is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe and parietal lobes to make those \emph{specific locations} touch-sensitive.  Focused on the internal features is the neon sign in Figure \ref{fig:neon}: the sign's output is based on the \emph{path} of the pipe.

@@

Pipe network topologies enable different interactions.

Splitting or mixing pipes offer flexibility in output.  If a maker wished to create a painting device, she might wish to have two system-side pipes feeding in primary-colored red and blue paints which mix in varying ratios, allowing their pigments to combine before purple exits from the device (see Figure \ref{fig:pipespace}).  Splitting can also be useful, for example if our maker wants red paint output in two locations from her painting device, she could have one system-side pipe, but split the pipe into two (see Figure \ref{fig:pipespace}).  \bjoern{??}

Star and tree topologies are extensions of the splitting and mixing primitives, encompassing multiple-in and multiple-out in contrast to single-in/single-out.  Using a star topology in which the pipes were filled with conductive paint, we created toys with several touch-sensitive areas, see Figure \ref{fig:toys}.

@@

``Gas'' comprises all compressible fluids.  Fluid pressure inside semi-closed pipes can create haptic feedback, or gases can be used as carriers for scents or fog.  Slyper, et al.,  in \cite{Slyper-pressure} engineered structures to change in air pressure when manipulated correctly (e.g., a spiral that changes pressure when twisted, but not when pressed); fluid pressure can also be used as an input.

Incompressible fluids (``liquids'') can perform many of the same interface tasks as gases.  One opportunity with liquids is to fill the interior of pipes with them and cap the ends.  In addition, one can use driable conductive fluids, such as copper paint, to coat the interior of pipes and allow them to function as arbitrarily-shaped wires.  This is especially helpful for the creation of a shared ground, or for creating single-wire capacitive interfaces amenable to sensing with SFCS \cite{Sato-touche}.

Pipes need not have hollow centers: in the case where routed pipes are filled with solid material---in particular, a solid material different from the model material---, interactions such as those in Printed Optics (\cite{Willis-printedoptics}) are possible.

Particulates, either printed in-place or inserted, can be of varying densities.  A single particle can be used for display.  Sparse particles in a stream of fluid can provide haptic feedback.  Dense particles in a semi-closed pipe capped with flexible surface material allow for jamming-based interactions at any point on the surface of an object \cite{Follmer-jamming}; when fluid pressure in the pipe is reduced, the trapped particles jam to form a solid mass the same shape as their enclosing flexible membrane.

Threadable inserted elements, such as electroluminescent (EL) wire or fiberoptic cables, are those that can be threaded through pipes post-printing.  This allows overcoming limitations of printers: for example, a Printed Optics-style interface can be created on an inexpensive consumer-grade 3D printer using pipes and inserted fiber optic cable.

>>>>

\section{The Design Space of Pipes}
By using different openings, designs, and topologies of pipes, and inserting different media into the completed pipes, inputs and outputs can be located at arbitrary points on the surface of a 3D print.  We describe the space below.

@@

Pipes can have four opening styles: open, return, semi-closed, and fully enclosed (see Figure \ref{fig:pipespace}).  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user; and a ``system side"---where other hardware such as electronics may be connected.  We refer to a pipe which is ``open on the user side'' to have its contents \emph{physically accessible} to a user \emph{at interaction time}.  ``Closed on the user side'' implies physically cut off from the user.  Similarly with the system side.  User or system manipulation of pipe contents are possible even without physical access, e.g., via soundwaves.

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/types.png}
\caption{Pipe openings: A shows a pair of open tubes filled with copper paint to power an LED.  B shows a return pipe with EL wire threaded through it.  C shows two semi-closed pipes capped by rubber membranes.  In D, a rabbit has a fully enclosed chamber to modify its weight. \bjoern{Great, but make figure  full column wide}}
\label{fig:openings}
\end{figure}

Open pipes originate from the system side and connect the user side, with both ends of the pipe open. This type of pipe may be used to create electronic circuitry.  An open pipe filled with conductive paint can power an LED (see Figure \ref{fig:openings}A).

A return pipe originates at the system and returns back to it, without offering physical access to the user.  By threading an electroluminescent (EL) wire through a clear return pipe, a maker can create a custom piece of neon art (see Figure \ref{fig:openings}B).

Semi-closed pipes are open on the system side, closed on the user side.  The closed interface may be fabricated from a malleable material, e.g., a rubber membrane, and attached to a fluid pump to create haptic feedback (see Figure \ref{fig:openings}C).

A fully enclosed pipe is closed to both the system and the user.  Fully enclosed pipes may be used as weight modifiers (e.g., for object identification), as in Figure \ref{fig:openings}D.  They can also contain for water or particles that would otherwise fall out.

\subsection{Design of Pipes}
Pipe design may emphasize either exterior connection points or interior paths.  Focused on the exterior features is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe and parietal lobes to make those \emph{specific locations} touch-sensitive. Alternatively, the neon sign's output in Figure \ref{fig:neon} is based on the \emph{path} of the pipe.

@@

Pipe network topologies enable different interactions.  Splitting or mixing pipes offer flexibility in output, allowing a single source to have multiple outputs or multiple sources to combine (see Figure \ref{fig:pipespace}).  Star and tree topologies extend the splitting and mixing primitives, encompassing multiple-in and multiple-out in contrast to single-in/single-out.

@@

``Gas'' comprises compressible fluids.  Fluid pressure inside semi-closed pipes can create haptic feedback, or gases can be used as carriers for scents or fog; fluid pressure can also be used as an input.

Incompressible fluids (``liquids'') perform similarly to gases.  Additionally, using driable conductive liquids like copper paint, pipes can become custom wires.

Pipes need not have hollow centers: in the case where routed pipes are filled with solid material---in particular, a solid material different from the enclosing model material---, interactions such as those in Printed Optics \cite{Willis-printedoptics} are possible.

Particulates, either printed in-place or inserted, can be of varying densities.  A single particle can be used for display.  Sparse particles in fluid can provide haptic feedback.

Threadable media, such as EL wire or fiberoptic cables, are those that can be threaded through pipes post-printing.  This empowers inexpensive printers: for example, a Printed Optics-style interface can be created on a consumer-grade 3D printer with pipes and fiber optic cable.

### 2014-04-01 paper/introduction.tex
shortening

\tovi{condense}
Makers, as well as professional designers, leverage 3D printers as tools for design work.  A wide array of objects, ranging from bicycle helmets to video game controllers, are now prototyped or even manufactured using these machines.  However, most devices fabricated by 3D printers are passive.  We propose a novel technique involving the removal of material from the interior of 3D models, prior to printing, to create pipes and other cavities.  These pipes can be filled, post-print, by a variety of media that enable input, display, and tactile feedback.  This {\em subtractive} approach is complementary to {\em additive} approaches that try to print different materials, such as conductors. While our approach requires manual assembly afterwards, it opens up a large useful design space on commonly available 3D printers, allowing makers to create objects like those in Figure \ref{fig:teaser}.

@@

Pipes allow makers to build with and upon existing input/output techniques, often integrating these methods into objects they were not originally designed for.   For example, to create a touch-sensitive toy as in Figure \ref{fig:teaser}, a maker can select several locations of interest on the brain.  After creating pipes to those locations and printing her object, she can fill the pipes with conductive paint.  For sensing, she need only connect a single wire to a shared out of all the tubes, and can distinguish them via Swept Frequency Capacitve Sensing \cite{Sato-touche}.  In this way, she has only a single active element (the microcontroller setup), but her otherwise-inert 3D printed object (the brain) has several sensing locations.  Other existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye} and Jamming User Interfaces \cite{Follmer-jamming}, can also be enhanced by the use of pipes: pipes allow makers utilizing such I/O strategies to locate input and output at arbitrary points on an object's surface.  Thus, we leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of I/O to arbitrary locations on a device.

>>>>

Makers, as well as professional designers, leverage 3D printers as tools for design work.  A wide array of objects, ranging from bicycle helmets to video game controllers, are now prototyped or even manufactured using these machines.  However, most devices fabricated by 3D printers are passive.  We propose a novel technique involving the removal of material from the interior of 3D models, prior to printing, to create pipes and other cavities.  These pipes can be filled, post-print, by a variety of media that enable input, display, and tactile feedback.  This {\em subtractive} approach is complementary to {\em additive} approaches that try to print different materials, such as conductors. While our approach requires manual assembly afterwards, it opens up a large useful design space on commonly available 3D printers, allowing makers to create objects like those in Figure \ref{fig:teaser}.

@@

Pipes allow makers to build with and upon existing input/output techniques, often integrating these methods into objects they were not originally designed for.   For example, to create a touch-sensitive toy as in Figure \ref{fig:teaser}, a maker can select several locations of interest on the brain.  After creating pipes to those locations and printing her object, she can fill the pipes with conductive paint.  For sensing, she need only connect a single wire to a shared out of all the tubes, and can distinguish them via Swept Frequency Capacitve Sensing \cite{Sato-touche}.  In this way, she has only a single active element (the microcontroller setup), but her otherwise-inert 3D printed object (the brain) has several sensing locations.  Other existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye} and Jamming User Interfaces \cite{Follmer-jamming}, can also be enhanced by the use of pipes: makers utilizing such I/O strategies can locate input and output at arbitrary points on an print's surface.  We leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of I/O to arbitrary locations on a device.
